---
title: "Lab_1_in_class"
author: "Danny Ettelson"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:

- basic review (data wrangling) 
- intro to spatial data in R with sf
- intro to Shiny apps

1. Attach Packages

```{r packages}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)


```

2. Review: reading exploring, wrangling, and visualizing data


a. Read in the data


```{r}

wine <- read_csv("wine_data.csv")

```

b. Wrangling

```{r Wrangling}

#select is columns, filter is rows
#can order columns with select 
#mutate can overwrite columns, best to just add
wine_new <- wine %>% 
  select(country,province,winery,region_1,points,price) %>%
  rename(state = province) %>% 
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% 
  mutate(ppd = points/price) %>% 
  arrange(-ppd)
  

```

c. Use group_by() and summarize() to find summary statistics by group

```{r group_by}

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE)
  )


```


d. A few gg-graphics

```{r graphs}
#histogram uses fill colors to group
wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
  facet_wrap( ~state, scales = "free")

wine_hists


```


new graph: scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it

```{r price_v_points}
#points is color not fill
#points use color, bars use fill and color as line color
wine_points <- ggplot(wine_new, aes(x=price,y=points)) +
  geom_point(aes(color = state), alpha = 0.5) +
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  facet_wrap(~state) +
  theme_light()


#there are packages that just have lots of themes to use. match publication style, etc.
wine_points


```

2. California Counties (Spatial Data) using sf

```{r}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016")

#plot(ca_counties)

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)

#full_join is safest, doesn't get rid of anything
ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome)

```


d. Use geom_sf to make a map in ggplot

```{r maps}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

#ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()

```












